\documentclass[11pt]{article}
% =====================================================================
%  I2S-pro family --- the four program-feature categories, explained.
%  Standalone explainer: one code example + the metric/rule + an
%  illustration matched to what each rule actually measures.
%  Compile: pdflatex i2s_pro_categories.tex   (needs tikz + pgfplots + listings)
% =====================================================================
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.16}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric}

\definecolor{winc}{RGB}{30,120,60}   % winner (I2S helps) green
\definecolor{losc}{RGB}{190,60,60}   % loser  (no I2S)   red
\definecolor{codebg}{RGB}{246,246,248}
\definecolor{kw}{RGB}{40,60,150}
\definecolor{cmt}{RGB}{120,120,120}
\definecolor{str}{RGB}{160,60,20}

\lstdefinestyle{c}{
  language=C, basicstyle=\ttfamily\small, backgroundcolor=\color{codebg},
  keywordstyle=\color{kw}\bfseries, commentstyle=\color{cmt}\itshape,
  stringstyle=\color{str}, numbers=none, frame=single, framerule=0pt,
  framesep=6pt, breaklines=true, showstringspaces=false, xleftmargin=4pt,
}

% A small "metric & rule" box used under every category.
\newcommand{\metricbox}[4]{%
  % #1 tool  #2 metric (plain meaning)  #3 verify rule  #4 real example
  \par\smallskip\noindent\fbox{\parbox{\dimexpr\linewidth-2\fboxsep-2\fboxrule}{%
  \footnotesize
  \textbf{Tool:}~\texttt{#1}\\[2pt]
  \textbf{Metric:}~#2\\[2pt]
  \textbf{Verify rule (the arbiter fires this, no LLM):}~\texttt{#3}\\[2pt]
  \textbf{Real example:}~#4}}\par\smallskip}

\title{\textbf{The I2S-pro family: four reasons input-to-state copying gets through}}
\author{}
\date{}

\begin{document}
\maketitle
\vspace{-3em}

\paragraph{Backdrop.} \emph{I2S-pro} collects the roadblocks where a fuzzer that
\emph{watches the program's comparisons and pastes the observed value back into
the input} (cmplog's input-to-state, I2S) gets through, while a byte-mutation
fuzzer stays stuck. Picture a lock whose key is lying right next to it: I2S
notices the key and uses it; plain fuzzing has to guess. All four categories
below share that technique --- they differ in \emph{why} it is the decisive one.
Counts are validated roadblocks ($n$).

% =====================================================================
\section*{1.\ \ Exact string / literal gate \hfill {\normalsize\texttt{i2s\_string\_literal\_substitution} ($n=185$)}}
The program demands an exact run of bytes: a magic header, a format tag
(\texttt{IDAT}, \texttt{sRGB}), a protocol token (\texttt{HTTP/}, \texttt{://}).
I2S sees that literal in the comparison and drops it straight into the input;
random mutation would essentially never spell it out. \emph{The key was lying
beside the lock.} This is the dominant case.

\begin{lstlisting}[style=c]
// lcms  cmspcs.c : _cmsLCMScolorSpace()
// dispatch on the 4-byte ICC color-space signature 'sig'
switch (sig) {
case cmsSigGrayData: return PT_GRAY;   // sig == 'GRAY'
case cmsSigCmykData: return PT_CMYK;   // <-- blocker (lcms/2246): sig == 'CMYK'
}
\end{lstlisting}

\noindent\textbf{How we know it is really I2S.} If I2S is planting the literal,
the cmplog corpus should be \emph{enriched} in those exact bytes at the gate
offset, relative to a naive corpus. We measure that enrichment as a log$_2$
frequency ratio; the rule asks for at least a $2\times$ enrichment.

\metricbox{operand\_enrichment}
{\texttt{signed\_target\_enrich} $=\log_2\!\dfrac{\text{cmplog freq}+\epsilon}{\text{naive freq}+\epsilon}$, $\epsilon{=}10^{-6}$ \ ($>0$ = enriched, $<0$ = depleted)}
{signed\_target\_enrich >= 1.0}
{lcms/2246 (\texttt{\_cmsLCMScolorSpace}) --- at the operand offset the 4-byte signature is \emph{absent} from the naive corpus (0\%) but reaches ${\approx}24\%$ in the cmplog corpus, so \texttt{signed\_target\_enrich}$=\log_2(0.24/\epsilon)=+17.9$; the decoy bytes are \emph{depleted} (\texttt{decoy\_enrich}$=-1.1$).}

\begin{center}
\begin{tikzpicture}[font=\footnotesize,
  fld/.style={draw,minimum height=6.5mm,minimum width=12mm,font=\scriptsize,inner sep=2pt},
  gate/.style={draw,very thick,fill=winc!15,minimum height=6.5mm,minimum width=18mm,font=\scriptsize},
  byte/.style={draw,minimum width=8.5mm,minimum height=6.5mm,font=\ttfamily\small},
  off/.style={font=\scriptsize\ttfamily,text=black!55}]
% --- ICC profile header: key fields at their real offsets ---
\node[fld] (size) {size};
\node[fld,right=1pt of size] (ver)  {version};
\node[fld,right=1pt of ver]  (dev)  {dev.\ class};
\node[gate,right=1pt of dev] (cs)   {colour space};
\node[fld,right=1pt of cs]   (pcs)  {PCS};
\node[fld,right=1pt of pcs]  (acsp) {\ttfamily 'acsp'};
\node[off,above=0pt of size]{0x00}; \node[off,above=0pt of ver]{0x08};
\node[off,above=0pt of dev]{0x0C};  \node[off,above=0pt of cs]{0x10};
\node[off,above=0pt of pcs]{0x14};  \node[off,above=0pt of acsp]{0x24};
\node[left=3pt of size,font=\scriptsize\itshape]{ICC header};
% --- enrichment callout on the gate field ---
\node[above=5mm of cs,font=\scriptsize,align=center,text=winc] (en)
     {\texttt{signed\_target\_enrich}\\$=\log_2(0.24/\epsilon)=\mathbf{+17.9}$};
\draw[->,winc,thick] (en) -- (cs.north);
% --- zoom the gate field into its 4 operand bytes ---
\node[byte,below=11mm of cs.south west,anchor=north west] (b0) {43};
\node[byte,right=0pt of b0] (b1) {4D};
\node[byte,right=0pt of b1] (b2) {59};
\node[byte,right=0pt of b2] (b3) {4B};
\node[font=\ttfamily\scriptsize,below=-1pt of b0]{'C'};
\node[font=\ttfamily\scriptsize,below=-1pt of b1]{'M'};
\node[font=\ttfamily\scriptsize,below=-1pt of b2]{'Y'};
\node[font=\ttfamily\scriptsize,below=-1pt of b3]{'K'};
\draw[dashed] (cs.south west) -- (b0.north west);
\draw[dashed] (cs.south east) -- (b3.north east);
\node[right=4mm of b3,align=left,font=\scriptsize]
     {exact gate operand \textbf{`CMYK'}\\(the data colour-space signature)};
% --- per-fuzzer frequency at this offset ---
\node[below=7mm of b0.south west,anchor=north west,align=left,font=\scriptsize] {%
  \textcolor{losc}{\textbf{naive}}\,:\ \ \,0\% of seeds carry \texttt{43 4D 59 4B} here (a different colour-space)\\
  \textcolor{winc}{\textbf{cmplog}}\,: $\approx$24\% carry it --- \texttt{I2SRandReplace} splices the logged \texttt{`CMYK'}};
\end{tikzpicture}\\[2pt]
{\footnotesize The gate operand of \texttt{lcms/2246} is the ICC \emph{data colour-space
signature} at offset \texttt{0x10}; \texttt{operand\_enrichment} measures how often each
fuzzer's corpus carries the exact bytes \emph{there}. Header field offsets are the ICC
format; the measured signal is \texttt{signed\_target\_enrich}$=+17.9$ (naive places the
literal in 0\% of its seeds), of which $\approx$24\% is the implied cmplog operand-offset rate.}
\end{center}

\paragraph{Sub-case: the literal \emph{floats} (no fixed offset).} Three of these
branches place the literal at a \emph{content-dependent} position rather than a
fixed file offset --- an XML entity name, a comment marker --- so the enrichment
test above cannot be anchored. The mechanism is identical (paste the exact
literal); only the operand's position differs, so we confirm them by the Layer-1
\mbox{cmplog\,$\succ$\,naive} advantage plus the \emph{absence} of any fixed-offset
signature (\texttt{gate\_signature == false}). Real example
(\texttt{libxml2}, \texttt{xmlGetPredefinedEntity}, \texttt{entities.c:307}):

\begin{lstlisting}[style=c]
switch (name[0]) {                       // name = a parsed entity reference
case 'a':                                // <-- blocker: needs the "amp"/"apos" name
    if (xmlStrEqual(name, "amp"))  return &xmlEntityAmp;
    if (xmlStrEqual(name, "apos")) return &xmlEntityApos;
    break;
default: return NULL;                    // naive lands here on all 479 of its calls
}
\end{lstlisting}
\noindent{\footnotesize cmplog takes \texttt{case 'a'} 63 times by splicing the
logged \texttt{"amp"}/\texttt{"apos"} operands into the input; naive never spells
them. The literal sits wherever the \texttt{\&\dots;} reference occurs, so there is
no fixed offset to measure --- hence the no-signature confirmation.}

% =====================================================================
\section*{2.\ \ Exact number / enum gate \hfill {\normalsize\texttt{i2s\_numeric\_tag\_substitution} ($n=39$)}}
Same move, but the thing matched is a small \emph{number} --- a version code, an
enum, a \texttt{switch}-case selector --- not a string. I2S substitutes the
precise value. We split it from category~1 only because ``it's a number'' vs
``it's a string'' matters when you later try to \emph{recognize} the gate; the
mechanism (paste the exact value) is identical.

\begin{lstlisting}[style=c]
// version/enum dispatch -- gate needs one exact constant
switch (header.version) {
  case 2:  parse_v2(...); break;    // <-- blocker: only the value 2 reaches it
  default: reject();
}
\end{lstlisting}

\noindent\textbf{How we know.} The cmplog corpus over-represents the \emph{target}
constant and \emph{under}-represents the competing ``decoy'' values, exactly as
deterministic substitution of one logged operand would predict.

\metricbox{operand\_enrichment}
{\texttt{signed\_target\_enrich} (target value, $\log_2$ ratio) and \texttt{decoy\_enrich} (the competing value)}
{signed\_target\_enrich >= 1.0 AND decoy\_enrich <= 0.0}
{bloaty/34, switch-case constant --- \texttt{signed\_target\_enrich}$=11.2$ (target piled up) while \texttt{decoy\_enrich}$=-3.7$ (wrong values drained).}

\begin{center}
\begin{tikzpicture}
\begin{axis}[
  width=11cm, height=3.4cm, xbar, xmin=-6, xmax=13,
  bar width=11pt, enlarge y limits=0.7,
  symbolic y coords={decoy value, target value},
  ytick=data, xtick={-3.7,0,11.2}, xticklabel style={font=\footnotesize},
  nodes near coords, every node near coord/.append style={font=\footnotesize},
  xlabel={\footnotesize log$_2$ enrichment in the cmplog corpus \ (right = piled up, left = drained)},
]
\addplot[fill=winc!70] coordinates {(11.2,target value)};
\addplot[fill=losc!60] coordinates {(-3.7,decoy value)};
\end{axis}
\end{tikzpicture}\\[-2pt]
{\footnotesize I2S concentrates the corpus on the one right number and steers it away from the wrong ones.}
\end{center}

% =====================================================================
\section*{3.\ \ Reassembling a structure to reach deeper \hfill {\normalsize\texttt{i2s\_structural\_assembly\_reach\_depth} ($n=8$)}}
Here it is not one literal but several \emph{interlocking} fields (an ELF section
table, a PNG \texttt{iTXt} chunk grammar, an XML scheme+attribute). I2S repairs
enough of those linked fields that the parser can finally march \emph{deeper}
into the input and arrive at the gate. The tell is not ``the right bytes are
present'' --- it is ``I2S's seeds drive execution much further in.''

\begin{lstlisting}[style=c]
// each field must validate before the parser advances to the next
if (read_len(&p)  != OK) return;     // field 1
if (read_type(&p) != OK) return;     // field 2  -- I2S repairs these...
if (read_kw(&p)   != OK) return;     // field 3
process_text(&p);                    // <-- blocker: only deep, well-formed seeds reach it
\end{lstlisting}

\noindent\textbf{How we know.} We run each side's corpus and count, per seed, how
many times the blocker line executes (its parse \emph{depth}) and whether the
seed takes the decisive side. The winner should reach the line far more.

\metricbox{depth\_reach}
{\texttt{depth\_lift} = winner / loser per-seed executions of the blocker line; \texttt{side\_lift} = how much more the winner takes the decisive branch; \texttt{winner\_deeper} (bool)}
{depth\_lift >= 2.0 AND side\_lift >= 3.0 AND winner\_deeper == true}
{bloaty/90 (\texttt{elf.cc:1279}, the ELF section-table loop) --- \texttt{depth\_lift}$=2375\times$, \texttt{side\_lift}$=2375$: carrying a valid ELF header, the I2S corpus drives the section loop three orders of magnitude deeper than naive.}

\begin{center}
\begin{tikzpicture}[font=\footnotesize,>=Latex,node distance=8mm]
\tikzstyle{f}=[draw,rounded corners,minimum width=20mm,minimum height=7mm]
\node[f] (a) {field 1: len};
\node[f,right=of a] (b) {field 2: type};
\node[f,right=of b] (c) {field 3: kw};
\node[f,right=of c,fill=winc!15] (d) {\textbf{GATE}};
\draw[->] (a)--(b); \draw[->] (b)--(c); \draw[->] (c)--(d);
% loser path stalls
\node[losc,below=5mm of a,align=center] (l) {loser (no I2S)\\stalls here};
\draw[->,losc,thick] (l) -- (a.south);
\draw[losc,thick,dashed] (a.south east) .. controls +(0.4,-0.5) and +(-0.4,-0.5) .. (b.south west) node[midway,below,losc]{$\times$};
% winner path reaches gate
\node[winc,above=5mm of c,align=center] (w) {winner (I2S repairs every field)};
\draw[->,winc,thick] (w.east) .. controls +(0.6,0.2) .. (d.north);
\end{tikzpicture}\\[2pt]
{\footnotesize I2S fixes the interlocking fields so the parser walks the whole chain to the gate; naive dies at the first one.}
\end{center}

% =====================================================================
\section*{4.\ \ Last-mile value precision \hfill {\normalsize\texttt{i2s\_operand\_value\_precision} ($n=6$)}}
The sneaky one. \emph{Both} sides \emph{reach} the gate line --- even the losing
fuzzers get there, sometimes more often. But only the I2S side lands the
\emph{exact} required value at the final step; the others get tantalisingly close
and never quite hit it. The edge is not \emph{getting} to the door, it is
\emph{turning the key precisely} at a door everyone is already standing at.

\begin{lstlisting}[style=c]
// everyone reaches this line; only the exact operand passes
if (strncmp(line, "Host:", 5) == 0)   // loser seeds hit "Host?", "Host:", ...
    route_host(line + 5);              // <-- blocker: needs the precise bytes
\end{lstlisting}

\noindent\textbf{How we know.} We measure how \emph{close} each side's nearest
seed gets to the target operand (a normalised distance, $0$ = exact hit). The
loser stalls at a positive residual; the winner closes that gap.

\metricbox{value\_distance\_reached}
{\texttt{cmp\_min\_distance} = loser's residual to the target ($>0$ = never lands it); \texttt{distance\_closure\_ratio} = fraction of the gap the winner closes; \texttt{distance\_gap}, \texttt{winner\_closer}}
{winner\_closer == true AND distance\_gap > 0 AND distance\_closure\_ratio >= 0.5}
{curl/129, operand \texttt{"Host:"} --- \texttt{distance\_gap}$=0.8$, \texttt{distance\_closure\_ratio}$=1.0$: the loser stalls $0.8$ away, the winner closes the whole gap.}

\begin{center}
\begin{tikzpicture}[font=\footnotesize,>=Latex]
\draw[->] (0,0) -- (10.5,0);
\foreach \x/\l in {0/0\,(exact hit), 5/0.5, 10/1.0\,(far)} \draw (\x,0.1)--(\x,-0.1) node[below]{\l};
\node[above] at (5,0.2) {distance to the target operand};
% loser marker at 0.8 -> x = 8
\filldraw[losc] (8,0) circle (3pt) node[above=4pt,losc,align=center]{loser (naive/vp)\\stalls at residual $0.8$};
% winner marker at 0
\filldraw[winc] (0,0) circle (3pt) node[above=4pt,winc,align=center]{winner (cmplog)\\lands it ($0$)};
\draw[->,winc,thick] (7.6,-0.45) -- (0.4,-0.45) node[midway,below]{closure $=1.0$ (closes the whole gap)};
\end{tikzpicture}
\end{center}

% =====================================================================
\section*{At a glance}
\begin{center}\small
\begin{tabular}{@{}p{3.6cm}clp{5.0cm}@{}}
\toprule
\textbf{Category} & \textbf{$n$} & \textbf{Tool / metric} & \textbf{One-line reason} \\
\midrule
string / literal gate & 185 & operand\_enrichment / \texttt{signed\_target\_enrich} & paste the exact bytes (incl.\ 3 floating-offset, \texttt{gate\_signature}=false) \\
numeric / enum gate & 39 & operand\_enrichment / \texttt{signed\_target\_enrich}+\texttt{decoy\_enrich} & paste the exact number \\
structural assembly depth & 8 & depth\_reach / \texttt{depth\_lift}, \texttt{side\_lift} & rebuild the structure to get there \\
operand value precision & 6 & value\_distance\_reached / \texttt{distance\_closure\_ratio} & get the value exactly right at the finish \\
\midrule
\textbf{I2S-pro total} & \textbf{238} & & \\
\bottomrule
\end{tabular}
\end{center}

\noindent\footnotesize Counts and example metrics regenerated from
\texttt{bench/dataset.jsonl}; verify rules are the deterministic arbiter rules
(no LLM in the decision).
\end{document}
