\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{enumitem}

\definecolor{kw}{RGB}{0,90,160}
\definecolor{cmt}{RGB}{110,110,110}
\definecolor{gatehl}{RGB}{200,30,30}

\lstdefinestyle{src}{
  basicstyle=\ttfamily\small,
  keywordstyle=\color{kw}\bfseries,
  commentstyle=\color{cmt}\itshape,
  numberstyle=\tiny\color{cmt},
  numbers=left, numbersep=6pt,
  frame=single, framesep=4pt,
  breaklines=true, showstringspaces=false,
  columns=fullflexible,
}
\lstdefinestyle{bytes}{
  basicstyle=\ttfamily\small,
  frame=single, framesep=4pt,
  breaklines=true, showstringspaces=false,
}

\newcommand{\ste}{\texttt{signed\_target\_enrich}}
\newcommand{\de}{\texttt{decoy\_enrich}}

\title{Three Ways Input-to-State Mutation \emph{Hurts}:\\
A Byte-Level Account of the I2S-Anti Mechanisms}
\author{BlockerAnalyzer benchmark --- I2S-anti family}
\date{2026-06-13}

\begin{document}
\maketitle

\begin{abstract}
In the BlockerAnalyzer benchmark, 41 decisive blocker branches are labeled
\emph{I2S-anti}: branches the input-to-state (I2S / cmplog) fuzzer
\emph{fails} to reach while a non-I2S arm (naive havoc or value-profile)
succeeds. All three I2S-anti sub-mechanisms share one measured signature ---
on the real campaign corpus, the cmplog arm is \emph{depleted} of the byte
value the winning arm reaches --- but they differ in \emph{why}. This note
defines the measurement at the byte level, then walks one specific branch per
mechanism with its real gate offset(s), byte values, and corpus statistics.
\end{abstract}

\section{The shared measurement: \texttt{operand\_enrichment}}

Every I2S-anti label is decided by one tool,
\texttt{i2s\_operand\_availability.py} (registry name
\texttt{operand\_enrichment}), run directly on the two campaign corpora.
The tool reads no source and assumes no mechanism; it measures bytes:

\begin{enumerate}[itemsep=2pt]
  \item \textbf{Gate signature.} Compare the branch's \emph{resolving} (W)
  seeds against its \emph{blocking} (L) seeds. The \emph{gate offsets} are the
  byte positions at which W and L seeds systematically disagree (consensus
  purity $\ge$ threshold). At each gate offset, the W-consensus byte is the
  \textbf{target} pattern, the L-consensus byte is the \textbf{decoy} pattern.
  \item \textbf{Corpus fractions.} For each fuzzer $F\in\{\textrm{cmplog},
  \textrm{naive}\}$, scan $F$'s saved corpus and compute
  \texttt{target\_frac}$_F$ (fraction whose head matches the full target
  pattern at all gate offsets) and \texttt{decoy\_frac}$_F$ (fraction matching
  the decoy pattern).
  \item \textbf{Signed enrichment.}
  \[
    \texttt{signed\_target\_enrich}=\log_2\!\frac{\texttt{target\_frac}_{\textrm{cmplog}}}
    {\texttt{target\_frac}_{\textrm{naive}}},\qquad
    \texttt{decoy\_enrich}=\log_2\!\frac{\texttt{decoy\_frac}_{\textrm{cmplog}}}
    {\texttt{decoy\_frac}_{\textrm{naive}}}.
  \]
\end{enumerate}

\noindent A value \ste{}$<0$ means cmplog's corpus contains the target byte
\emph{less} often than naive's --- I2S is steering bytes \emph{away} from the
value that resolves the branch. All three labels require
\ste{}$<-0.3$; they are separated by the decisive \emph{shape} (which fuzzers
win/lose) and by \emph{what} the decoy is.

\begin{table}[h]\centering
\caption{The three I2S-anti mechanisms at a glance. Shape code is over
$[\text{cmplog},\text{vp},\text{vpc},\text{naive}]$ with W=win / L=loss /
\_=non-decisive.}
\label{tab:overview}
\begin{tabular}{@{}llll@{}}
\toprule
label & shape & decoy is\ldots & rule \\
\midrule
\texttt{i2s\_decoy\_substitution\_target\_depletion} & \texttt{LW\_W} &
  a competing real operand & \ste{}$<-0.3$ \emph{and} gate present \\
\texttt{i2s\_anti\_structural\_byte\_corruption}     & \texttt{LWWW} &
  --- (blind overwrite) & \ste{}$<-0.3$ \\
\texttt{i2s\_decoy\_substitution\_overfit}           & \texttt{LWLW} &
  a structural ``stay-valid'' literal & \ste{}$<-0.3$ \\
\bottomrule
\end{tabular}
\end{table}

% ======================================================================
\section{Mechanism 1 --- \texttt{target\_depletion}: a rival operand wins the harvest}
\label{sec:depletion}
\textbf{Example branch: \texttt{libpng\_3977}} (validated, $n=23$ for this label).

The gate is a single byte. \texttt{operand\_enrichment} found
\texttt{gate\_offsets} $=[236]$, \texttt{gate\_width} $=1$. The branch is the
\texttt{cHRM} (chromaticity) chunk handler comparing a parsed fixed-point value
to the error sentinel:

\begin{lstlisting}[style=src,language=C,
  caption={\texttt{/src/libpng/pngrutil.c:1283}, \texttt{OSS\_FUZZ\_png\_handle\_cHRM}.
  The gate byte at input offset 236 feeds \texttt{xy.bluex}.},label={lst:libpng}]
   if (xy.redx   == PNG_FIXED_ERROR ||
       xy.redy   == PNG_FIXED_ERROR ||
       xy.greenx == PNG_FIXED_ERROR ||
       xy.greeny == PNG_FIXED_ERROR ||
       xy.bluex  == PNG_FIXED_ERROR ||   /* <-- decisive branch 3977 */
       xy.bluey  == PNG_FIXED_ERROR)
      /* error path: chunk rejected */
\end{lstlisting}

At offset 236 the resolving seeds carry the \textbf{target byte
\texttt{0x00}}; the blocking seeds carry the \textbf{decoy byte \texttt{0x65}}
(ASCII \texttt{'e'}). The decoy is not random: it is the exponent character of
ASCII numeric literals (\texttt{"1.0e0"}, \texttt{"65.535"}) that cmplog
harvests from \emph{other} string/number comparisons in the PNG parser and
splices in via \texttt{I2SRandReplace}:

\begin{lstlisting}[style=bytes,caption={Hex window over bytes 228--244 around
  the gate byte (in brackets at offset 236). Top: a resolving (W) seed. Bottom:
  a blocking (L) cmplog seed --- note the embedded ASCII float 1.0e0 whose
  e (0x65) lands exactly on the gate.},label={lst:libpngbytes}]
W (resolving, naive)  : 02 02 02 02 02 75 30 00 [00] ea 60 20 00 79 98 00 00
L (blocking,  cmplog) : 2f 62 61 72 00 31 2e 30 [65] 30 00 36 35 2e 35 33 35
                                       (  "1.0e0"  )         ( "65.535" )
\end{lstlisting}

\begin{table}[h]\centering
\caption{\texttt{libpng\_3977} operand\_enrichment metrics
(\texttt{--sample 8000 --head 256}). cmplog is depleted of the target
\texttt{0x00} and enriched in the decoy \texttt{0x65}.}
\label{tab:libpng}
\begin{tabular}{@{}lrr@{}}
\toprule
quantity & cmplog & naive \\
\midrule
corpus size scanned                 & 4561   & 4447   \\
\texttt{target\_frac} (byte \texttt{0x00} @236) & 0.0886 & 0.1565 \\
\texttt{decoy\_frac}  (byte \texttt{0x65} @236) & 0.6290 & 0.3929 \\
\midrule
\multicolumn{3}{@{}l}{$\ste=\log_2(0.0886/0.1565)=\mathbf{-0.821}$ \quad(target depleted)}\\
\multicolumn{3}{@{}l}{$\de=\log_2(0.6290/0.3929)=\mathbf{+0.679}$ \quad(decoy enriched)}\\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{Reading:} a real, abundant competing operand (the textual
\texttt{0x65}) exists at the same byte position. cmplog logs it and
RandReplaces it in, \emph{pinning} 63\% of its corpus to the decoy and dropping
the target \texttt{0x00} to 8.9\% (vs.\ 15.7\% for naive, which never harvests
the decoy and drifts onto the target freely). The extra rule clause
\texttt{skip != 'no\_gate\_signature'} guarantees a genuine two-operand
competition, not noise.

% ======================================================================
\section{Mechanism 2 --- \texttt{byte\_corruption}: blind RandReplace clobbers an assembled byte}
\label{sec:corruption}
\textbf{Example branch: \texttt{harfbuzz\_5831}} (validated; server-a target).

Shape \texttt{LWWW}: cmplog is the \emph{sole} loser (naive, value\_profile,
and vpc all win) --- the clean fingerprint of ``only the I2S mutator breaks
this.'' Here there is no clever rival operand. Naive havoc \emph{assembles} the
correct structural byte; cmplog's \texttt{I2SRandReplace} blindly overwrites
that exact position with a logged value, \emph{destroying} the byte rather than
redirecting it to a single dominant decoy.

\begin{table}[h]\centering
\caption{\texttt{harfbuzz\_5831} operand\_enrichment metrics. The target byte
is depleted in cmplog ($\ste<0$); decoy fractions are small in absolute terms
(diffuse corruption), not one dominant rival.}
\label{tab:hb5831}
\begin{tabular}{@{}lrr@{}}
\toprule
quantity & cmplog & naive \\
\midrule
\texttt{target\_frac} & 0.2113 & 0.3799 \\
\texttt{decoy\_frac}  & 0.0293 & 0.0030 \\
\midrule
\multicolumn{3}{@{}l}{$\ste=\mathbf{-0.882}$ \quad(target byte clobbered out of cmplog's corpus)}\\
\multicolumn{3}{@{}l}{$\de=\mathbf{+2.64}$ \quad(weak, low-absolute --- diffuse, not a single decoy)}\\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{Reading:} the structural byte is present in 38\% of naive's
corpus but only 21\% of cmplog's --- I2S keeps overwriting it. Contrast with
Mechanism~1: there the decoy is a \emph{single abundant} value
(\texttt{decoy\_frac}$=0.63$); here the decoy mass is small and scattered
(\texttt{decoy\_frac}$=0.029$), the signature of \emph{corruption} (random
clobber) rather than \emph{competition} (capture by one rival).

\paragraph{Byte-level proof (server a).} Mechanism~2 is \emph{defined} by the
absence of a single decoy, so --- unlike M1 and M3 --- there is no clean
``target $\to$ one decoy'' byte to exhibit; the evidence \emph{is} the
diffuseness. Recomputed on harfbuzz's corpus, the gate is wide and scattered
(\texttt{gate\_width}$=29$, offsets 71--187) --- itself a tell that no single
byte is decisive. The assembled structural byte is depleted in cmplog and
present in naive (per-offset median \texttt{target\_frac} $0.20$ vs.\ $0.34$,
reproducing the assignment's $0.21/0.38$), but the bytes cmplog writes there do
\emph{not} concentrate on any replacement:

\begin{lstlisting}[style=bytes,caption={Byte distribution at gate offset 71
  (assembled target \texttt{0x00}) over the 8000-input corpora. naive peaks on
  the target; cmplog is spread over 211 distinct values with no dominant
  substitute --- the fingerprint of blind corruption, not a decoy.},
  label={lst:hb5831bytes}]
naive  @71 : 0x00=0.39  0xff=0.03  0x05=0.03  0x01=0.03  ...   (236 distinct)
cmplog @71 : 0x00=0.32  0x10=0.12  0x01=0.07  0x04=0.06  ...   (211 distinct)
\end{lstlisting}

\noindent Across the 29 gate offsets the \emph{largest single non-target byte}
in cmplog reaches only $\approx 0.12$ (median), versus the one dominant decoy of
M1 ($0.63$) and M3 ($0.51$): nothing to capture, nothing re-asserted, only
scatter. Seed length corroborates the assembly story --- cmplog's blocking seeds
are short (10--61\,B, one $405$\,B outlier) while the resolving naive seeds run
60--188\,B: cmplog \emph{truncates / fails to assemble} the record rather than
substituting one wrong value into a complete one. This is why M2, alone among
the three, has no clean two-row decoy window: corruption is diffuse by
construction.

\paragraph{Caveat: why naive resolves despite its own noise, and what that
implies for the label.} A fair objection: naive's corpus is \emph{also} noisy at
the gate --- at offset 71 it has \emph{more} distinct bytes (236 vs.\ 211) and
only a slightly higher target share ($0.39$ vs.\ $0.32$). So ``naive is
cleaner'' is \emph{not} why it wins. Resolution does not require a low-noise
corpus; it requires \emph{one} input that assembles the structure and is
\emph{kept} by coverage feedback. The discriminator here is structural
completeness, visible in seed length: naive's resolving seeds run 60--188\,B
while cmplog's blocking seeds are 10--61\,B (one $405$\,B outlier) --- and
\emph{all} 29 gate offsets lie at byte $\ge 71$, so eight of cmplog's nine
blocking seeds are shorter than the \emph{first} gate offset. Naive's plain
havoc occasionally completes the record and coverage latches the long input;
cmplog's \texttt{I2SRandReplace} overwrites bytes mid-assembly, so its corpus
stays short and broken and never latches a resolving input. The per-byte scatter
is the residue of that repeated clobber, not the cause.

\noindent This refines the label rather than contradicting it. The winner/loser
median seed-length ratio is $\approx 1.7$ for \texttt{harfbuzz\_5831}
(cmplog truncated) --- the \emph{opposite} sign from the M3 overfit case below,
where cmplog's blocking seeds are \emph{longer} (ratio $\approx 0.4$, bloated
``stay-valid'' inputs). \texttt{operand\_enrichment} cannot see this: it reports
\ste{}$<0$ for both. So \texttt{harfbuzz\_5831} sits between pure in-place
byte-corruption and an \emph{assembly/truncation failure} --- the byte is
depleted because the record never completes, not (only) because a complete
record is clobbered. A \texttt{size\_lift} (winner-vs-loser seed size) clause
would subtype these honestly, but the present I2S-anti member set is too small
($n\!=\!8$ for this label) to fit such a split without over-anchoring on a single
branch, so it is left as a noted refinement rather than a new mechanism.

% ======================================================================
\section{Mechanism 3 --- \texttt{overfit}: structural over-fitting that starves a \emph{semantic} (exotic-script) gate}
\label{sec:overfit}
\textbf{Example branch: \texttt{harfbuzz\_5323}} (validated; server-a target).

Shape \texttt{LWLW}: both I2S carriers (cmplog and the value-profile-cmplog arm)
lose; naive and value-profile win. \textbf{Unlike M1/M2 the decisive gate is not
an input byte at all} --- \texttt{harfbuzz\_5323} is
\texttt{hb\_script\_get\_horizontal\_direction:594}, \texttt{case
HB\_SCRIPT\_THAANA}, so the branch resolves iff the \emph{decoded text carries
Thaana-script codepoints} (\texttt{U+0780--07BF}), a \emph{semantic} property
harfbuzz computes from the text, with no fixed input offset. cmplog \emph{does
reach} the branch (it is a genuine blocker, hit-status $2/8/0$ --- reaches the
\texttt{switch} every trial, returns \texttt{LTR} for its non-RTL text); what it
fails to do is \emph{produce} exotic-script text, because its \texttt{I2SRandReplace}
keeps splicing harvested structural literals (font table tags, magic) that bias
the corpus toward structurally-valid, ASCII/Latin fonts. It ``overfits'' to
structural validity and thereby starves the script the gate needs.

\begin{table}[h]\centering
\caption{\texttt{harfbuzz\_5323} operand\_enrichment metrics --- the extreme
case. cmplog's corpus is half the structural decoy; naive's has almost none.}
\label{tab:hb5323}
\begin{tabular}{@{}lrr@{}}
\toprule
quantity & cmplog & naive \\
\midrule
\texttt{target\_frac} & 0.0816 & 0.5083 \\
\texttt{decoy\_frac}  & 0.5075 & 0.0004 \\
\midrule
\multicolumn{3}{@{}l}{$\ste=\mathbf{-2.64}$ \quad(target free-byte present in 51\% of naive vs 8\% of cmplog)}\\
\multicolumn{3}{@{}l}{$\de=\mathbf{+10.4}$ \quad(structural literal dominates cmplog, near-absent in naive)}\\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{Reading:} this is the strongest anti-signal in the family.
cmplog spends 51\% of its corpus re-asserting the structural literal
(\texttt{decoy\_frac}$_{\textrm{cmplog}}=0.5075$ vs.\ $0.0004$ for naive). But
note what the byte statistic actually is here: a \emph{correlate}, not the gate
(see the correction below). Where Mechanism~1 is \emph{two operands competing}
and Mechanism~2 is \emph{one byte being corrupted}, Mechanism~3 is \emph{a
validity trap}: I2S ``succeeds'' at keeping the input structurally valid and
thereby starves the exotic-script text the semantic gate requires.

\paragraph{Byte-level proof (server a).} Recomputing the gate directly on
harfbuzz's corpus identifies the structural decoy concretely: it is the
\textbf{first OpenType table tag} --- the 4-byte field at input offsets
\texttt{[12:16]}, immediately after the 12-byte \texttt{sfnt} header
(\texttt{sfntVersion}(4)+\texttt{numTables}(2)+\texttt{searchRange}(2)+%
\texttt{entrySelector}(2)+\texttt{rangeShift}(2)). In cmplog's blocking seeds
those four bytes spell a \emph{registered} tag --- \texttt{GPOS},
\texttt{GSUB}, or \texttt{MATH}; in the resolving (naive) seeds the same
position is unstructured:

\begin{lstlisting}[style=bytes,caption={Hex window over bytes 8--23. The 4-byte
  table tag (brackets, offsets 12--15) is the literal \texttt{GPOS}
  (\texttt{47 50 4f 53}) in the cmplog seed and random in the naive seed. The
  following 8 bytes are the table record's \texttt{checkSum} and
  \texttt{offset}.},label={lst:hb5323bytes}]
W (resolving, naive)  : ea 00 00 ff [00 1b 00 00] 63 61 72 66 15 03 00 0e
L (blocking,  cmplog) : 21 20 1e 20 [47 50 4f 53] 00 01 00 0d 00 00 00 10
                                    (   "GPOS"  )  (checkSum)  (tbl off)
\end{lstlisting}

\noindent At corpus scale the tag is overwhelmingly an I2S artifact. Of
cmplog's 8000-input sample, \textbf{50.3\%} carry exactly \texttt{GPOS} at
offsets 12--15 and \textbf{82.9\%} carry \emph{some} printable table tag there;
for naive the figures are \textbf{0.0\%} and \textbf{6.1\%}. This reproduces the
tool's \texttt{decoy\_frac} ($0.508$ vs.\ $0.0004$) at the byte level. Crucially,
the ``literal'' is a \emph{tag field}, not one fixed constant: cmplog harvests
whichever registered tag the OpenType \texttt{CMP} checks expose
(\texttt{GPOS}/\texttt{GSUB}/\texttt{MATH}) and re-asserts it to keep the font's
table directory well-formed --- precisely why the signature is ``structural''
rather than a single magic number.

\paragraph{Correction: \texttt{GPOS} is a \emph{correlate}, not the decisive
operand.} The byte tool finds the bytes that most \emph{discriminate} the two
corpora; here those are the font table tags. But the branch's source
(\texttt{hb-common.cc:594}) is the \texttt{HB\_SCRIPT\_THAANA} case of a
script-direction \texttt{switch} --- the decisive condition is the text's
\emph{script}, which has no fixed input offset. Coverage proves it: in the
resolving (naive) per-role dump the THAANA case at line 594 is hit \textbf{34}
times and every earlier RTL case (\texttt{ARABIC}/\texttt{HEBREW}/\texttt{SYRIAC})
\textbf{0}, so the resolving text is unambiguously Thaana; in the blocking
(cmplog) dump line 594 is \textbf{0} --- not because cmplog fails to reach the
\texttt{switch} (it does, every trial) but because its text is never Thaana, so
it returns \texttt{LTR}. \texttt{GPOS} co-varies with blocking only because
cmplog's I2S simultaneously (i) injects font structure and (ii) suppresses exotic
scripts; the tag is the visible side-effect, not the cause. \emph{General lesson:
for a \textbf{semantic-property} gate the byte-frequency tool necessarily reports
a structural correlate, never the operand --- the label fires for the right
reason (cmplog's corpus is depleted of what resolves) but the byte-level "decoy"
is a proxy.}

\paragraph{The \texttt{overfit} cluster is heterogeneous.} Five of the seven
\texttt{overfit}-labelled branches are harfbuzz \emph{exotic-Unicode} gates
(\texttt{hb\_script\_get\_horizontal\_direction} $\times2$, combining-class
recategorisation, Hebrew compose, Hangul compose) --- one coherent
``exotic-script starvation'' mechanism. The remaining two
(\texttt{openthread\_13300/13302}, \texttt{LinkedList::PopAfter}/\texttt{Find}
over the address-resolver cache) share only the \texttt{LWLW} \emph{shape}: they
are a stateful data-structure depth blocker with \emph{no} decoy
(\texttt{decoy\_enrich}$\approx0.08$), and cmplog reaches them ($2/8/0$) exactly
as it reaches the harfbuzz members. No byte or reachability metric separates the
two mechanisms (the \texttt{decoy\_enrich} of the five harfbuzz members ranges
$-3.3$ to $+10.4$, straddling the OpenThread value), so the pair is correctly
left \emph{inconclusive} (guardrail G3) rather than forced under a label whose
mechanism they do not share --- a worked example of the deterministic benchmark's
ceiling: mechanisms that differ only \emph{semantically} cannot be auto-split.

\paragraph{A note on the gate tool's offsets.} Run at the default consensus
purity ($0.6$) the tool reports a 3-offset, \emph{non-contiguous} gate
(\texttt{[11,14,15]}, decoy bytes \texttt{" OS"}) --- a \emph{fragment} of the
tag: it keeps only offsets where \emph{both} W and L seeds reach consensus, and
the heterogeneous naive side has none at offsets 12--13. The full 4-byte tag is
recovered by dumping the cmplog (L) seeds or the corpus-fraction test above.

% ======================================================================
\section{Side-by-side}
\begin{table}[h]\centering
\caption{The discriminating contrast. All three have \ste{}$<0$ (target
depleted in cmplog); the \emph{decoy} statistics separate them.}
\label{tab:contrast}
\begin{tabular}{@{}lrrr@{}}
\toprule
 & \textbf{M1 depletion} & \textbf{M2 corruption} & \textbf{M3 overfit} \\
 & \texttt{libpng\_3977} & \texttt{harfbuzz\_5831} & \texttt{harfbuzz\_5323} \\
\midrule
shape                         & \texttt{LW\_W} & \texttt{LWWW} & \texttt{LWLW} \\
\ste{}                        & $-0.82$ & $-0.88$ & $-2.64$ \\
\de{}                         & $+0.68$ & $+2.64$ & $+10.4$ \\
\texttt{decoy\_frac} cmplog   & 0.629 & 0.029 & 0.508 \\
\texttt{decoy\_frac} naive    & 0.393 & 0.003 & 0.0004 \\
decoy character               & one abundant rival & diffuse / none & structural literal (\emph{correlate}) \\
winner gate                   & input byte (operand) & assembled byte & \emph{semantic} script property (no byte) \\
gate is causal byte?          & yes & yes & no --- byte ``decoy'' is a proxy \\
one-line                      & rival wins harvest & byte clobbered & validity trap starves a script \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Provenance note.} \texttt{libpng\_3977} is a server-sB target;
its gate offset, target/decoy bytes, and seed windows above were extracted
directly from the on-disk campaign corpus. \texttt{harfbuzz\_5831} and
\texttt{harfbuzz\_5323} are server-a targets; their metric values are taken
verbatim from \texttt{step5b\_new\_v3/*/assignments\_s4.json}, and their
byte-level evidence (Listings~\ref{lst:hb5831bytes} and~\ref{lst:hb5323bytes})
was extracted directly from server a's on-disk campaign corpus
(\texttt{cmplog}/\texttt{naive} queues, \texttt{--sample 8000}). The slight
differences between the reproduced per-offset fractions and the assignment's
stored values are sampling noise from the random corpus draw.

\end{document}
